---
title: "Multiple Comparisons with a Plan"
author: "Your Group's Names Here!"
date: "Due: November 5, 2019"
output: html_document
---

# White-Nose Syndrome

For this (Halloween) lab we will consider data on white-nose syndrome (WNS). 
White-nose syndrome (WNS) is a disease that affects hibernating bats and is 
caused by a fungus, Pseudogymnoascus destructans. Sometimes this fungus looks 
like a white fuzz on bats’ faces, which is how the disease got its name. The 
fungus grows in cold, dark and damp places. It attacks the bare skin of bats 
while they’re hibernating in a relatively inactive state. As it grows, it causes
changes in bats that make them become active more than usual and burn up fat 
they need to survive the winter. Bats with white-nose syndrome may do strange 
things like fly outside in the daytime in the winter. 

White-nose syndrome has killed millions of bats in North America. At some sites,
90 to 100 percent of bats have died. Several species are affected, with the 
hardest-hit being the northern long-eared bat, little brown bat, and tricolored
bat. Other species, like the Virginia big-eared bat, have been found with white
nose, but they don’t show signs of being sick with white-nose syndrome. There is
no cure for white-nose syndrome, but scientists from all over the world are 
working together to study the disease, how it spreads and infects bats and what 
we can do to control it. 

These data were collected by Ecologists in Indiana, interested in studying the 
spread of WNS throughout Indiana. WNS is suspected to have infected Indiana
caves in the winter of 2010-2011, so these researchers plan to use long-term 
monitoring data to examine changes in a summer population that may be 
attributable to WNS effects during winter.  


# Data 

These researchers collected capture data in central Indiana on 10 repeatedly 
netted sites consistent across all reproductive periods of the bats. They then 
used a cokriging model to estimate the probability of WNS infection for their 
capture data. 

```{r, message = FALSE}
library(tidyverse)

wns <- read_csv("data/white_nose_Pettit_Okeefe.csv") 

```


__1. Load in the data and run the code below to create a summary table of the 
number of sampling days at each site each year.__ 
__What do you notice about the measurements for sited D, D1, G, and I?__


```{r}
wns %>% 
  group_by(year, Site.Name) %>% 
  summarize(count = n()) %>% 
  spread(key = year, value = count, fill = 0)
```


## Combining Levels into One 

Sites D and D1 are not different, and neither are sites G and I. But they were 
coded differently in the data. To combine these levels into one level, we will 
use the `recode()` function in `dplyr`!  

```{r}

```





# Findings 

The little brown bat Myotis lucifugus, the
Indiana bat M. sodalis, and the tri-colored bat Perimyotis subflavus suffered 12–80% declines; whereas,
the big brown bat Eptesicus fuscus, the eastern red bat Lasiurus borealis, and the evening bat Nycticeius
humeralis showed 11.5–50.5% increases. We caught more nonreproductive adult females and
postlactating females when WNS exposure probabilities were high, suggesting that WNS is influencing
reproductive success of affected species. We conclude that, in Indiana, WNS is causing species-specific
declines and may have caused the local extinction of M. lucifugus. Furthermore, WNS-affected species
appear to be losing pups or forgoing pregnancy. Ongoing long-term monitoring studies, especially those
focusing on reproductive success, are needed to measure the ultimate impacts of WNS.


# References 

Pettit, Joseph L.; O'Keefe, Joy M. (2017), Data from: Impacts of white-nose 
syndrome observed during long-term monitoring of a midwestern bat community, 
Dataset, https://doi.org/10.5061/dryad.cf273


