---
title: "Multiple Comparisons with a Plan"
author: "Your Group's Names Here!"
date: "Due: November 5, 2019"
output: html_document
---

# White-Nose Syndrome

For this (Halloween) lab we will consider data on white-nose syndrome (WNS). 
White-nose syndrome (WNS) is a disease that affects hibernating bats and is 
caused by a fungus, Pseudogymnoascus destructans. Sometimes this fungus looks 
like a white fuzz on bats’ faces, which is how the disease got its name. The 
fungus grows in cold, dark and damp places. It attacks the bare skin of bats 
while they’re hibernating in a relatively inactive state. As it grows, it causes
changes in bats that make them become active more than usual and burn up fat 
they need to survive the winter. Bats with white-nose syndrome may do strange 
things like fly outside in the daytime in the winter. 

White-nose syndrome has killed millions of bats in North America. At some sites,
90 to 100 percent of bats have died. Several species are affected, with the 
hardest-hit being the northern long-eared bat, little brown bat, and tricolored
bat. Other species, like the Virginia big-eared bat, have been found with white
nose, but they don’t show signs of being sick with white-nose syndrome. There is
no cure for white-nose syndrome, but scientists from all over the world are 
working together to study the disease, how it spreads and infects bats and what 
we can do to control it. 

These data were collected by Ecologists in Indiana, interested in studying the 
spread of WNS throughout Indiana. WNS is suspected to have infected Indiana
caves in the winter of 2010-2011, so these researchers plan to use long-term 
monitoring data to examine changes in a summer population that may be 
attributable to WNS effects during winter.  


# Data 

These researchers collected capture data in central Indiana on 10 repeatedly 
netted sites consistent across all reproductive periods of the bats. They then 
used a cokriging model to estimate the probability of WNS infection for their 
capture data. 

```{r, message = FALSE}
library(tidyverse)

wns <- read_csv("data/white_nose_Pettit_Okeefe.csv") 

```


__1. Load in the data and run the code below to create a summary table of the 
number of sampling days at each site each year.__ 
__What do you notice about the measurements for sited D, D1, G, and I?__


```{r}
wns %>% 
  group_by(year, Site.Name) %>% 
  summarize(count = n()) %>% 
  spread(key = year, value = count, fill = 0)
```


## Combining Levels into One 

Sites D and D1 are not different, and neither are sites G and I. But they were 
coded differently in the data. To combine these levels into one level, we will 
revisit the `case_when()` function in `dplyr`!  

__2. Write the code to:__ 
* __change `Site.Name` D1 to D__ 
* __change `Site.Name` G to I__  

```{r}
wns <- wns %>% 
  mutate(Site.Name = case_when(Site.Name == "D1" ~ "D", 
                               Site.Name == "G" ~ "I", 
                               TRUE ~ Site.Name)
         )
```


## Reordering Levels 

As we all know, the default ordering of characters in `R` is alphabetical. 
However, sometimes we wish to have a specific ordering to our characters (e.g. 
Monday, Tuesday, Wednesday, etc.). The default ordering for the bat species is 
not what we would like to use. Instead, we would like the levels to be: 

EPFU, PESU, NYHU, LABO, MYLU, MYSE, MYSO

To reorder variable levels, we will use the `fct_relevel()` function from the 
`forcats` package. The package `forcats` (the name is an anagram of factors!) is
included in the `tidyverse`, so we don't need to load in a new package.  


```{r}
wns <- wns %>% 
  mutate(species_tidy = fct_relevel(species, 
                                    "EPFU", 
                                    "PESU", 
                                    "NYHU", 
                                    "LABO", 
                                    "MYLU", 
                                    "MYSE", 
                                    "MYSO")
         )
```



## EDA  

__3. Using `ggplot`, create a visualization of the number of bats captured for  
each species.__ 
__Verify, is the order of the species what you specified in the code above?__ 


```{r}
wns %>%  
  ggplot(aes(x = species, y = bats)) + 
  geom_violin(aes(fill = species)) + 
  facet_wrap(~year)
```


### Faceting 

Almost always it is useful to consider **more** than two variables in a data 
visualization. To add a third variable into our plot, we could make the same 
plot as above, but produce a subplot for each year. To do this, we use the 
`facet_wrap()` function in `ggplot2` to add a facet layer to our plot. 

__4. Add a facet to the plot you made above. What new information do you see?__ 



# Comparisons 

These researchers investigated differences in the number of captures for each 
species, controlling for the site bats were captured on.  

__5. Fit a two-way ANOVA using `aov` modeling the number of bats captured as 
a function of the study site and bat species.__ 
__Next, fit a multiple comparison test for differences in the number of bats 
captured per species, using `TukeyHSD`.__ 
__Finally, present the results in a tidy table!__  

```{r}
library(broom)

bats_aov <- aov(bats ~ Site.Name + species, data = wns)

bats_thsd <- TukeyHSD(bats_aov, "species", ordered = TRUE)

tidy(bats_thsd)
```


__6.__ How many comparisons were made? How many had p-values less than 0.05? How
many would you expect to have a p-value less than 0.05 due to random chance?__ 




## Changing Data Type 

When we imported the data, `year` variable was stored as a numeric variable, but
the researchers analyzed these data using `year` as a `factor`. 

To convert a variable from numeric to a character we use `R`'s built-in 
`as.character()` function. 

```{r}
wns <- wns %>% 
  mutate(yearF = as.character(year))
  
```

__7. Confirm that `yearF` is considered to be a character data type and not a 
numeric data type.__  



# More Comparisons  

Researchers were also interested in investigating the differences in the number
of captures for each species, controlling for the year were captured in.  

__8. Fit a two-way ANOVA using `aov` modeling the number of bats captured as 
a function of the year captured and bat species.__ 
__Next, fit a multiple comparison test for differences in the number of bats 
captured per year, using `TukeyHSD`.__ 
__Finally, present the results in a tidy table!__  

```{r}
year_aov <- aov(bats ~ yearF + species, data = wns)

year_thsd <- TukeyHSD(year_aov, "yearF", ordered = TRUE)

tidy(year_thsd)
```


__9.__ How many comparisons were made? How many had p-values less than 0.05? How
many would you expect to have a p-value less than 0.05 due to random chance?__ 




# Reporting Findings 

Suppose the report that these researchers produced for the Journal of Fish and 
Wildlife Management solely reported the following results:  

* Our research found that big brown bat (Eptesicus fuscus) had a significantly 
lower capture rate than all other species of bats captured. Furthermore, the 
number of bats captured in 2012 was significantly higher than all other years of
the study.*  


__10. What are some issues with reporting the statistical analysis in this way?__ 





__If you are struggling to find issues, the following blog may be helpful:__ 
https://www.nih.gov/about-nih/what-we-do/science-health-public-trust/perspectives/science-health-public-trust/tips-communicating-statistical-significance


# References 

Pettit, Joseph L.; O'Keefe, Joy M. (2017), Data from: Impacts of white-nose 
syndrome observed during long-term monitoring of a midwestern bat community, 
Dataset, https://doi.org/10.5061/dryad.cf273


